\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-27 paper/main.tex
paper: added todos
paper: check biblio, add nocite, add a bit to intro for importance of data cleaning, remove an old reference to acs
paper fix: added option to hyperref package to avoid error with urls split across lines
paper: fixed general typos, added clearer todos
paper: added some preliminary experiments
paper: pass over motivating examples section
paper: partial pass over algorithm section
Fix whitespace in query
paper: working on optimality proof
paper: cleaned up proof, first pass
paper: intro feedback from martin
paper: abstract (phillip feedback)

\documentclass{vldb}
\usepackage{balance}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[activate={true,nocompatibility},final,kerning=true,spacing=true,factor=1100,stretch=10,shrink=10]{microtype}
\usepackage[subtle]{savetrees}
\usepackage{amsfonts,amssymb,amsmath}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[style=numeric,firstinits=true]{biblatex}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{hyperref} % load after other packages
\usepackage{graphicx} % includegraphics
\usepackage{listings} % code samples with syntax hl
\usepackage{stmaryrd}
\usepackage{qtree} % for simple tree diagrams
\usepackage{subcaption}
\addbibresource{references.bib}

@@

\newcommand{\ProjectName}{ImputeDB}

@@

\author{
  Jos\'e Cambronero \\
  \texttt{jcamsan@mit.edu}
  \and
  John K. Feser \\
  \texttt{feser@csail.mit.edu}
  \and
  Micah Smith \\
  \texttt{micahs@mit.edu}}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\todo}[1]{{\color{red}\textbf{TODO:} #1}}

\begin{document}

@@

\begin{abstract}
Missing values abound in data analysis and often present a serious challenge for users, who
are forced to pick between removing tuples with missing values or creating a clean duplicate of their data
by  applying a potentially expensive imputation strategy. We show that it is possible
to incorporate imputation into a cost-based query optimizer, performing imputation on the fly
for each query. This allows users to immediately explore
their data, while the system picks the optimal placement of imputation operations based
on the specific query issued and a parameter that the user selects for evaluating the tradeoff between data
quality and runtime performance. We incorporate the imputation operators into the relational algebra, allowing
for traditional optimizations to take place. Our experiments evaluate this approach on two real-world survey-based datasets
from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}, showing that the 
optimizer generates query plans that execute in a fraction of the time that it takes to impute values at the base
tables, making exploratory data analysis using our system feasible. Furthermore, we show that the query results
from on-the-fly imputation are comparable to those generated with the standard full-table imputation approach.
\end{abstract}

\input{introduction}
\input{examples}
\input{relatedwork}
\input{algorithm}
\input{experiments}
\input{conclusion}

\balance
\printbibliography

>>>>

\documentclass{vldb}
\usepackage{balance}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[activate={true,nocompatibility},final,kerning=true,spacing=true,factor=1100,stretch=10,shrink=10]{microtype}
\usepackage[subtle]{savetrees}
\usepackage{amsfonts,amssymb,amsmath}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[style=numeric,firstinits=true]{biblatex}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{graphicx} % includegraphics
\usepackage{listings} % code samples with syntax hl
\usepackage{stmaryrd}
\usepackage{qtree} % for simple tree diagrams
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{hyperref} % load after other packages
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{cleveref}

\addbibresource{references.bib}

@@

\newcommand{\ProjectName}{ImputeDB}
\newcommand{\doms}[2]{#1 \succ #2}

% definitions/theorems etc
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{case}{Case}

% new section names for cref
\Crefname{query}{Query}{Query}

@@

\author{
  Jos\'e Cambronero \\
  \texttt{jcamsan@csail.mit.edu}
  \and
  John K. Feser \\
  \texttt{feser@csail.mit.edu}
  \and
  Micah J. Smith \\
  \texttt{micahs@mit.edu}}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\todo}[1]{{\color{red}\textbf{TODO:} #1}}
\newcommand{\todobox}[2]{{\color{red}\fbox{\parbox{\columnwidth}{\textbf{TODO:} #1\\#2}}}}
\begin{document}

@@

\begin{abstract}
Missing values abound in data analysis and often present a serious challenge for usability of data. Users
are forced to pick between removing tuples with missing values or creating a clean duplicate of their data
by  applying a potentially-expensive imputation strategy. We show that it is possible
to incorporate imputation into a cost-based query optimizer, performing imputation on the fly
for each query. This allows users to immediately explore
their data, while the system picks the optimal placement of imputation operations. This placement is based
on the specific query issued and a parameter that the user selects for evaluating the tradeoff between data
quality and runtime performance. We incorporate the imputation operators into the relational algebra, allowing
for traditional optimizations to take place. Our experiments evaluate this approach on two real-world survey-based datasets
from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}. Our results show that the optimizer
generates query plans that execute between \todo{add [low x - high x] range} as fast as imputing at the base
tables and then executing queries, making exploratory data analysis using our system feasible. Furthermore, we show that the query results
from on-the-fly imputation have Symmetric-Mean-Absolute-Percentage-Error between \todo{add SMAPE range}  
when compared to standard full-table imputation approach.
\end{abstract}

\input{introduction}
\input{examples}
\input{relatedwork}
\input{algorithm}
\input{experiments}
\input{conclusion}
\balance
\nocite{*}
\printbibliography
